### 2012-09-29 doc/prelim/document/chapters/introduction/introduction.tex
updating section headers

\chapter{Introduction}
\label{ch:introduction}
In many fields of engineering and physics, linear and nonlinear
problems are a primary focus of study. Recent focus on multiple
physics systems in the nuclear reactor modeling and simulation
community adds a new level of complexity to common nonlinear systems
as solution strategies change when they are coupled to other problems
\citep{u.s._department_of_energy_casl:_2011}. Furthermore, a desire
for predictive simulations to enhance the safety and performance of
nuclear systems creates a need for extremely high fidelity
computations to be performed for these coupled systems as a means to
capture effects not modeled by coarser methods.

In order to achieve this high fidelity, state-of-the-art computing
facilities must be leveraged in a way that is both efficient and
considerate of hardware-related issues. As scientific computing moves
towards exascale facilities with machines of $O(1,000,000)$ cores
already coming online, new algorithms to solve these complex problems
must be developed to leverage this new hardware
\citep{kogge_using_2011}. Issues such as resiliency to node failure,
limited growth of memory available per node, and scaling to large
numbers of cores will be pertinent to robust algorithms aimed at this
new hardware. Considering these issues, this preliminary report
proposes the development of a massively parallel stochastic method for
linear problems and a novel stochastic method to advance solution
techniques for nonlinear problems.

@@

In addition to linear solver advancements, nonlinear solvers may also
benefit from a general and parallel MCSA scheme. In the nuclear
engineering community, nonlinear problems are often addressed by
either linearizing the problem or building a segregated scheme and
using traditionaly iterative or direct methods to solve the resulting
system \citep{pletcher_computational_1997}. In the mathematics
community, various Newton methods have been popular
\citep{kelley_iterative_1995}. Recently, Jacobian-Free Newton-Krylov
(JFNK) schemes \citep{knoll_jacobian-free_2004} have been utilized in
multiple physics architectures and advanced single physics codes
\citep{gaston_parallel_2009}. The benefits of JFNK schemes are that
the Jacobian is never formed, simplifying the implementation, and a
Krylov solver is leveraged (typically GMRES or Conjugate Gradient),
providing excellent convergence properties for well-conditioned and
well-scaled systems. However, there are two potential drawbacks to
these methods for high fidelity predictive simulations: (1) the
Jacobian is approximated by a first-order differencing method on the
order of machine precision such that this error can grow beyond that
of those in a fine-grained system \citep{kelly_iterative_1995} and (2)
for systems that are not symmetric positive-definite (which will be
the case for most multiphysics systems and certainly for most
preconditioned systems) the Krylov subspace generated by the GMRES
solver may become prohibitively large
\citep{knoll_newton-krylov_1995}. To address these issues, this thesis
proposes a new and novel method for nonlinear systems based on the
MCSA method.

>>>>

\chapter{Introduction}
\label{ch:introduction}
In many fields of engineering and physics, linear and nonlinear
problems are a primary focus of study. Recent focus on multiple
physics systems in the nuclear reactor modeling and simulation
community adds a new level of complexity to common nonlinear systems
as solution strategies change when they are coupled to other problems
\citep{u.s._department_of_energy_casl_2011}. Furthermore, a desire
for predictive simulations to enhance the safety and performance of
nuclear systems creates a need for extremely high fidelity
computations to be performed for these coupled systems as a means to
capture effects not modeled by coarser methods.

In order to achieve this high fidelity, state-of-the-art computing
\index{facilities} must be leveraged in a way that is both efficient and
considerate of hardware-related issues. As scientific computing moves
towards exascale facilities with machines of $O(1,000,000)$ cores
already coming online, new algorithms to solve these complex problems
must be developed to leverage this new hardware
\citep{kogge_using_2011}. Issues such as resiliency to node failure,
limited growth of memory available per node, and scaling to large
numbers of cores will be pertinent to robust algorithms aimed at this
new hardware. Considering these issues, this preliminary report
proposes the development of a massively parallel stochastic method for
linear problems and a novel stochastic method to advance solution
techniques for nonlinear problems.

@@

In addition to linear solver advancements, nonlinear solvers may also
benefit from a general and parallel MCSA scheme. In the nuclear
engineering community, nonlinear problems are often addressed by
either linearizing the problem or building a segregated scheme and
using traditionaly iterative or direct methods to solve the resulting
system \citep{pletcher_computational_1997}. In the mathematics
community, various Newton methods have been popular
\citep{kelley_iterative_1995}. Recently, Jacobian-Free Newton-Krylov
(JFNK) schemes \citep{knoll_jacobian-free_2004} have been utilized in
multiple physics architectures and advanced single physics codes
\citep{gaston_parallel_2009}. The benefits of JFNK schemes are that
the Jacobian is never formed, simplifying the implementation, and a
Krylov solver is leveraged (typically GMRES or Conjugate Gradient),
providing excellent convergence properties for well-conditioned and
well-scaled systems. However, there are two potential drawbacks to
these methods for high fidelity predictive simulations: (1) the
Jacobian is approximated by a first-order differencing method on the
order of machine precision such that this error can grow beyond that
of those in a fine-grained system \citep{kelley_iterative_1995} and
(2) for systems that are not symmetric positive-definite (which will
be the case for most multiphysics systems and certainly for most
preconditioned systems) the Krylov subspace generated by the GMRES
solver may become prohibitively large
\citep{knoll_newton-krylov_1995}. To address these issues, this thesis
proposes a new and novel method for nonlinear systems based on the
MCSA method.

### 2012-09-29 doc/prelim/document/chapters/research_proposal/research_proposal.tex
intro updates
updating section headers

\section{Proposed Challenge Problem}
\label{sec:challenge_problem}
Our numerical experiments will provide us with a set of metrics giving
best-performance stochastic solver parameters for problems of
different types. As an additional bonus, these experiments will let us
further develop new parallel implementations of the stochastic methods
in order to improve their performance and scalability. With these
advancements in hand, we look forward towards applying them to
large-scale problems in reactor physics. The Consortium for Advanced
Simulation of Light Water Reactors (CASL) is a Department of Energy
Modeling and Simulation Hub focused on predictive multiphysics
modeling of nuclear reactors
\citep{u.s._depertment_of_energy_casl:_2011}. Their effors are focused
on enabling nuclear reactor analysis and development using today's
supercomputing technology to achieve higher power uprates, higher fuel
burnup, and predictive accident scenario analysis. As a means of
driving their research and development, CASL has worked with the
nuclear industry to identify challenge problems of critical importance
to achieving these goals.

To meet some of the high fidelity modeling requirements of these
challenge problems, the Drekar multiphysics code is being developed at
Sandia National Laboratories for fluid flow and conjugate heat
transfer problems \citep{pawlowski_drekar::cfd_2012}. Drekar is being
used to analyze complex problems including the grid-to-rod-fretting
(GTRF) phenomena observed in light water reactor fuel rods. GTRF
occurs due to the turbulent flow of water through the subchannels of
the reactor core. This turbulence causes the nuclear fuel pins to
vibrate and the fuel cladding to wear against structures in the
reactor, causing localized hot spots and ultimately cladding
failure. Reducing these instances of failure and understanding how
they occur through predictive multiple physics simulation of fluid
flow and heat transfer will permit more robust fuel assembly
designs. In addition, the turbulence studies that Drekar performs will
allow for improved flow patterns to be generated in the assembly
subchannels, allowing for better heat transfer between the nuclear
fuel and the water, ultimately allowing for a more efficient reactor.

>>>>

\section{Proposed Challenge Problem}
\label{sec:challenge_problem}
Our numerical experiments will provide us with a set of metrics giving
best-performance stochastic solver parameters for problems of
different types. As an additional bonus, these experiments will let us
further develop new parallel implementations of the stochastic methods
in order to improve their performance and scalability. With these
advancements in hand, we look forward towards applying them to
large-scale problems in reactor physics. The Consortium for Advanced
Simulation of Light Water Reactors (CASL) is a Department of Energy
Modeling and Simulation Hub focused on predictive multiphysics
modeling of nuclear reactors
\citep{u.s._department_of_energy_casl_2011}. Their effors are focused
on enabling nuclear reactor analysis and development using today's
supercomputing technology to achieve higher power uprates, higher fuel
burnup, and predictive accident scenario analysis. As a means of
driving their research and development, CASL has worked with the
nuclear industry to identify challenge problems of critical importance
to achieving these goals.

To meet some of the high fidelity modeling requirements of these
challenge problems, the Drekar multiphysics code is being developed at
Sandia National Laboratories for fluid flow and conjugate heat
transfer problems \citep{pawlowski_drekar_2012}. Drekar is being
used to analyze complex problems including the grid-to-rod-fretting
(GTRF) phenomena observed in light water reactor fuel rods. GTRF
occurs due to the turbulent flow of water through the subchannels of
the reactor core. This turbulence causes the nuclear fuel pins to
vibrate and the fuel cladding to wear against structures in the
reactor, causing localized hot spots and ultimately cladding
failure. Reducing these instances of failure and understanding how
they occur through predictive multiple physics simulation of fluid
flow and heat transfer will permit more robust fuel assembly
designs. In addition, the turbulence studies that Drekar performs will
allow for improved flow patterns to be generated in the assembly
subchannels, allowing for better heat transfer between the nuclear
fuel and the water, ultimately allowing for a more efficient reactor.
